---
title: "reticulate-r-py-Nreproducibility-demo"
format: html
execute: 
  echo: true     # Show code
  warning: false # Hide warnings 
  message: false # Hide messages
  error: false   # Continue on error
---

## Some Introduction

#### What is Quarto
Quarto is an open-source scientific and technical publishing system that lets you combine narrative text with code to create reproducible and elegantly formatted output.

It is similar to Rmarkdown, a polished extension to it. A big feature of Quarto is that it also supports languages apart from R, like say Python.

Quarto documents have two main sections: 

- (i) the YAML header, where we specify document-wide properties such as the output format, and 
- (ii) the content, which can include text, images, code, and more.

For this demo, we directly jump into the project without describing further about YAML and reticulate. You can find these additional details from Nicola's blog post ([Combining R and Python with {reticulate} and Quarto](https://nrennie.rbind.io/blog/combining-r-and-python-with-reticulate-and-quarto/)).

Lets start with reading the data. You can download the data from the [#TidyTuesday repository](https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-08-24/readme.md).

```{r}
#| label: read-data
#| cache: true
lemurs <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-08-24/lemur_data.csv')
```

```{r}

library(reticulate)
# Explicitely set the path to use Conda environment
use_condaenv("retic_demo", required = TRUE)
```

```{r}
#| label: wrangling
#| echo: true
#| message: false
#| output-location: slide

library(dplyr)
library(knitr)

lemur_data <- lemurs |> 
  filter(taxon == "ECOL",
          sex == "M",
          age_category == "adult") |> 
  select(c(age_at_wt_mo, weight_g)) |> 
  rename(Age = age_at_wt_mo,
          Weight = weight_g)

kable(head(lemur_data)) 
```

Now, we can add a Python code block to fit a model:

```{r}
library(reticulate)
py_install("statsmodels")
```

```{r}
# In R, install poetry via reticulate
# library(reticulate)
# py_install("poetry", pip = TRUE)

# Or from command line (run in terminal/CMD)
# pip install poetry
```


Then in R
```{r setup, include = FALSE}

library(reticulate)

# use_poetry("E:/My Drive/MSc DS Essex/Anaesthesia/Rpractise/my-ci-cd-projects/retic-rpy-demo")


# Find your Poetry virtual environment path
# Run this in your terminal to get the path:
# poetry env info --path

# Set the Python interpreter to Poetry's venv
# use_python("E:/My Drive/MSc DS Essex/Anaesthesia/Rpractise/my-ci-cd-projects/retic-rpy-demo/python.exe", required = TRUE)

# Or use virtualenv path directly
# Sys.setenv(RETICULATE_PYTHON = "path/to/poetry/venv/python.exe")
```

```{python}
#| label: modelling
#| echo: true
#| message: false
lemur_data_py = r.lemur_data

import statsmodels.api as sm
y = lemur_data_py[["Weight"]]
x = lemur_data_py[["Age"]]
x = sm.add_constant(x)
mod = sm.OLS(y, x).fit()
lemur_data_py["Predicted"] = mod.predict(x)
lemur_data_py["Residuals"] = mod.resid
```

The key part here is the r. on line 1(or line 4 if you are looking at a .qmd file) - this tells {reticulate} to look in the R code for an object called lemur_data. You can think of this line as importing your data from R to Python. Other than that, it’s pretty standard Python code. Here, we’ve kept it very simple and fitted a linear model:

Obviously, R has the capabilities to fit a linear model, and for this simple example there’s no need for us to use Python. However, this illustrates how you might approach the problem, if you find yourself needing to use a particular library or model that you can only find in Python.

After we’ve fitted any type of model, we need to check the output. For linear regression, one of the most common methods of model inspection is to plot and analyse the residuals. Residuals tell us how far away a point is from the regression line. For linear models, the main assumption is that the errors are independent and normally distributed. If our data fits that assumption, we should see that:

the residuals are symmetric above and below 0;
we don’t see any patterns in the residuals as the predicted values increase;
most of the points are close to 0, and there are fewer points with higher magnitude of residual.

Whilst you can plot residuals in Python quite easily, in my opinion it’s hard to beat the quality of graphics you can achieve with {ggplot2} in R. So we include a final R code block to plot the residuals:

```{r}
#| label: plotting
#| echo: true
#| output-location: slide
#| message: false
#| fig-align: center
#| fig-alt: "Scatter plot of predicted and residual values for the fitted linear model."
library(reticulate)
library(ggplot2)
lemur_residuals <- py$lemur_data_py
ggplot(data = lemur_residuals,
       mapping = aes(x = Predicted,
                     y = Residuals)) +
  geom_point(colour = "#2F4F4F") +
  geom_hline(yintercept = 0,
             colour = "red") +
  theme(panel.background = element_rect(fill = "#eaf2f2",
                                        colour = "#eaf2f2"),
        plot.background = element_rect(fill = "#eaf2f2",
                                       colour = "#eaf2f2"))
```
















